A spatio-temporal analysis of the environmental correlates of COVID-19 incidence in Spain

### 2020-06-08 Environmental-Correlates-of-COVID19-Spain/Environmental-Correlates-of-COVID19-Spain.Rmd
RELEASE. Updated preprint with names of authors and added doi information to README.Rmd

---
title: A spatio-temporal analysis of the environmental correlates of COVID-19 incidence in Spain
author:
  - name: Author A
    email: author.a@example.com
    affiliation: Some School
    footnote: 1
  - name: Author B
    email: author.b@example.com
    affiliation: Some Institute
address:
  - code: Some School
    address: School of the Things
  - code: Some Institute
    address: Institute of Everything
  - code: Departamento de Economia
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
  The novel SARS-CoV2 has disrupted health systems and the economy and public health interventions to slow its spread have been costly. How and when to ease restrictions to movement hinges in part on whether SARS-CoV2 will display seasonality due to variations in temperature, humidity, and hours of sunshine. Here, we address this question by means of a spatio-temporal analysis in Spain of the incidence of  COVID-19, the disease caused by the virus. Use of spatial Seemingly Unrelated Regressions (SUR) allows us to model the incidence of reported cases of the disease per 100,000 population as an interregional contagion process, in addition to a function of temperature, humidity, and sunshine. In the analysis we also control for GDP per capita, percentage of older adults in the population, population density, and presence of mass transit systems. The results support the hypothesis that incidence of the disease is lower at higher temperatures and higher levels of humidity. Sunshine, in contrast, displays a positive association with incidence of the disease. Our control variables also yield interesting insights. Higher incidence is associated with higher GDP per capita and presence of mass transit systems in the province; in contrast, population density and percentage of older adults display negative associations with incidence of COVID-19.    
journal: "Geographical Analysis"
date: "`r Sys.Date()`"
bibliography: References.bib
#linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

>>>>

---
title: A spatio-temporal analysis of the environmental correlates of COVID-19 incidence in Spain
author:
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    footnote: 1
  - name: Fernando A. Lopez
    email: fernando.lopez@upct.es
    affiliation: Universidad Politecnica de Cartagena
  - name: Tatiane Menezes
    email: tatiane.menezes@ufpe.br
    affiliation: Departamento de Economia
  - name: Renata Cavalcanti
    email: renata.vcsantos@gmail.com
    affiliation: Nucleo de Pesquisa
  - name: Maira Galdino da Rocha Pitta
    email: mgrpitta@ufpe.br
    affiliation: Nucleo de Pesquisa
address:
  - code: McMaster University
    address: School of Geography and Earth Sciences, McMaster University, 1281 Main St W, Hamilton, ON, L8S 4K1, Canada
  - code: Universidad Politecnica de Cartagena
    address: Departamento de Metodos Cuantitativos, Ciencias Juridicas, y Lenguas Modernas, Universidad Politecnica de Cartagena, Calle Real Numero 3, 30201, Cartagena, Murcia, Spain
  - code: Departamento de Economia
    address: Departamento de Economia, Universidade Federal de Pernambuco, Av dos Economistas, s/n - Cidade Universitária, Recife - PE, 50670-901, Brasil  
  - code: Nucleo de Pesquisa
    address: Núcleo de Pesquisa em Inovação Terapêutica NUPIT / UFPE, Av. Prof. Moraes Rego, 1235 - Cidade Universitária, Recife, PE, CEP 50670-901, Brazil  
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
  The novel SARS-CoV2 has disrupted health systems and the economy and public health interventions to slow its spread have been costly. How and when to ease restrictions to movement hinges in part on whether SARS-CoV2 will display seasonality due to variations in temperature, humidity, and hours of sunshine. Here, we address this question by means of a spatio-temporal analysis in Spain of the incidence of  COVID-19, the disease caused by the virus. Use of spatial Seemingly Unrelated Regressions (SUR) allows us to model the incidence of reported cases of the disease per 100,000 population as an interregional contagion process, in addition to a function of temperature, humidity, and sunshine. In the analysis we also control for GDP per capita, percentage of older adults in the population, population density, and presence of mass transit systems. The results support the hypothesis that incidence of the disease is lower at higher temperatures and higher levels of humidity. Sunshine, in contrast, displays a positive association with incidence of the disease. Our control variables also yield interesting insights. Higher incidence is associated with higher GDP per capita and presence of mass transit systems in the province; in contrast, population density and percentage of older adults display negative associations with incidence of COVID-19.    
journal: "Geographical Analysis"
date: "`r Sys.Date()`"
bibliography: References.bib
#linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

### 2020-06-08 Environmental-Correlates-of-COVID19-Spain/Environmental-Correlates-of-COVID19-Spain.tex
RELEASE. Updated preprint with names of authors and added doi information to README.Rmd

  \title{A spatio-temporal analysis of the environmental correlates of COVID-19
incidence in Spain}
    \author[Some School]{Author A\corref{1}}
   \ead{author.a@example.com} 
    \author[Some Institute]{Author B}
   \ead{author.b@example.com} 
      \address[Some School]{School of the Things}
    \address[Some Institute]{Institute of Everything}
    \address[Departamento de Economia]{}
      \cortext[1]{Corresponding Author}

>>>>

  \title{A spatio-temporal analysis of the environmental correlates of COVID-19
incidence in Spain}
    \author[McMaster University]{Antonio Paez\corref{1}}
   \ead{paezha@mcmaster.ca} 
    \author[Universidad Politecnica de Cartagena]{Fernando A. Lopez}
   \ead{fernando.lopez@upct.es} 
    \author[Departamento de Economia]{Tatiane Menezes}
   \ead{tatiane.menezes@ufpe.br} 
    \author[Nucleo de Pesquisa]{Renata Cavalcanti}
   \ead{renata.vcsantos@gmail.com} 
    \author[Nucleo de Pesquisa]{Maira Galdino da Rocha Pitta}
   \ead{mgrpitta@ufpe.br} 
      \address[McMaster University]{School of Geography and Earth Sciences, McMaster University, 1281 Main
St W, Hamilton, ON, L8S 4K1, Canada}
    \address[Universidad Politecnica de Cartagena]{Departamento de Metodos Cuantitativos, Ciencias Juridicas, y Lenguas
Modernas, Universidad Politecnica de Cartagena, Calle Real Numero 3,
30201, Cartagena, Murcia, Spain}
    \address[Departamento de Economia]{Departamento de Economia, Universidade Federal de Pernambuco, Av dos
Economistas, s/n - Cidade Universitária, Recife - PE, 50670-901, Brasil}
    \address[Nucleo de Pesquisa]{Núcleo de Pesquisa em Inovação Terapêutica NUPIT / UFPE, Av.
Prof.~Moraes Rego, 1235 - Cidade Universitária, Recife, PE, CEP
50670-901, Brazil}
      \cortext[1]{Corresponding Author}
